---
output: github_document
editor_options: 
  chunk_output_type: console
bibliography: "inst/metapower.bib"
csl: "inst/apa.csl"
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(badger)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# metapoweR <img src = 'man/figures/metapower_sticker.png' align = "right" height = "180" />

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/metapower)](https://CRAN.R-project.org/package=metapower)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://www.tidyverse.org/lifecycle/#stable)

<!-- badges: end -->

The primary goal of metapower is to compute statistical power for meta-analyses. Currently, metapower has the following functionality:

Computation of statistical power for:

1. Summary main effects sizes
2. Test of homogeneity for between-group variance (for Random-effects models). 
3. Test of homogeneity for within-study variance
4. Categorical moderator analyses

metapower can currently handle the following designs and effect sizes:

4. Standardized mean difference: Cohen's *d*
5. Correlation between two continuous variables: Correlation Coefficient (via Fisher's r-to-z transformation)
6. Probability of Success/Failure: Odds Ratio

## Installation

```{r include = FALSE}
#You can install the released version of metapower from [CRAN](https://CRAN.R-project.org) with:
#
#
#install.packages("metapower")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jasonwgriffin/metapower")
```
## Example

```{r, dpi = 300,fig.height=7, fig.width=7}
library(metapower)
my_power <- mpower(effect_size = .25, sample_size = 20, k = 30, es_type = "d")
print(my_power)
power_plot(my_power)
```

See Vignette "Using metapower" for more information

## References

All mathematical calculations are derived from @hedges2004, @bornstein2009, and @pigott2012.

<div id="refs"></div>

## Issues

If you encounter a clear bug, please file a minimal reproducible example on [github](https://github.com/jasonwgriffin/metapower/issues).



