---
title: "Computing power with metapower"
author: "Jason W. Griffin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using-metapower}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  #collapse = TRUE,
  #comment = "#>"
)
library(metapower)
```



The primary goal of metapower is to compute statistical power for meta-analyses. Currently, metapower is designed to compute statistical power for the following under fixed- and random-effects models:

1. Mean effect size difference between groups (e.g., Cohen's *d*)
2. Test of homogenity for between-study variance
3. Categorical moderator analyses of the mean effect size

# Example 1: Computing power to detect mean difference effect size

Suppose that we plan to meta-analyze all published findings to compute a summary effect size estimate for the group difference between typically developing individuals and indivudals with autism on a measure of face reocognition ability. In order to plan the study accordingly, we must choose plausible values for the following:

1. Expected effect size
2. Expected sample size per group
3. Expected number of studies
4. Expected degree of heterogenity (only for Random-effects model)

...*for our meta-analysis of face recognition deficits in autism*

1. We expect that face recognition deficits in ASD are small (Cohen's d = 0.25)
2. Sample sizes in autism research are generally small. We expect the average group size to be 20.
3. Face recognition is frequently studied in autism; therefore, we expect to find 50 studies.
4. Autism is notoriously heterogenous. We expect large heterogenity between-studies.

To do this with `metapower`, we use the core function `mpower()`

```{r}
my_power <- mpower(effect_size = .25, 
                   sample_size = 20, 
                   k = 50, 
                   hg = "large", 
                   es_type = "d",
                   model = "random", 
                   test_type = "two-tailed")
```

Note that we specify this a random-effects model (`model = "random`). 
For fixed-effects model, use `model = "fixed"`.

```{r}
print(my_power)
```

The first part of the output shows the expected input values, where the main results are shown in the bottom portion, mainly, `Estimated Power`. Under this set of values, our power to detect a mean difference is `r round(my_power$power*100,2)`%.



```{r fig.height=4, fig.width=7}
power_plot(my_power)
```

